Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-04-14 thesis.qmd
final

---
title: "Analysis of observational data, to calculate the Delayed-Tau model parameters"
author: "Dimitrios Papachristopoulos"
fontfamily: fontenc
format: 
  pdf: default  
  # arxiv-pdf:
  #    keep-tex: true    
  #    link-citations: true
  #    link-bibliography: true
  #    runninghead: ""
toc: true
toc-expand: true
toc-depth: 3
number-sections: true
appendix-style: plain
bibliography: ./My_Library.bib

@@

abstract: |
  The Star Formation History (SFH) of a galaxy can offer many insights not only for the evolution and the future of the galaxy, but also for the evolution of the Universe. This is why there are various theoretical models trying to describe the SFH of galaxies. One of those models is the Delayed-Tau model, which approximates the star formation of galaxies as a function with a rising Star Formation Rate (SFR) at the beginning, until it reaches a peak at a time $\tau$, unique for each galaxy, and then it drops at an exponential rate.

@@

  **Keywords:** Galaxies, Galaxy Evolution, Star Formation History (SFH), Star Formation Rate (SFR), Delayed-$\tau$, Local Cosmological Volume, Lilly-Madau Plot, Redshift, Newton-Ramphson, Markov Chain Monte Carlo (MCMC).
---

@@

# Star Formation

The SFH of a galaxy describes the evolution of its star formation rate over time. By selecting an appropriate model for SFH, we can analyze stellar production, predict periods of active or quiescent star formation, and determine when SFR stabilizes.

@@

<!-- {{<pagebreak>}} -->

>>>>

---
  title: "Analysis of observational data, to calculate the Delayed-Tau model parameters"
  subtitle: "Undergraduate thesis"
  author: 
  - name: Dimitrios Papachristopoulos
    role: sdf
    affiliations:
      - id: undgdt
        department: ID number 1110 2018 00142 
        name: Undergraduate student
  - name: Despina Hatzidimitriou 
    affiliations:
      - id: sprvsr
        name: Supervising professor 
        department: National and Kapodistrian University of Athens
  fontfamily: fontenc
  mainfont: "Times New Roman"
  format: 
    titlepage-pdf: 
      titlepage: vline
      titlepage-logo: "template/uoa_new_logo_red.png"
      titlepage-footer: | # text
        National and Kapodistrian University of Athens\
        School of Science\
        Faculty of Physics\
        Department of Astrophysics, Astronomy and Mechanics\
        April 2025
      author-style: plain
      author-sep: "\ --\ "
      number-sections: true  
      documentclass: scrartcl
      classoption: ["oneside", "open=any"]

    #pdf: default  
    # arxiv-pdf:
    #    keep-tex: true    
    #    link-citations: true
    #    link-bibliography: true
    #    runninghead: ""
  toc: true
  toc-expand: true
  toc-depth: 3
  number-sections: true
  appendix-style: plain
  bibliography: ./My_Library.bib

@@

---
{{<pagebreak>}}

# Abstract

## English

The Star Formation History (SFH) of a galaxy can offer many insights not only for the evolution and the future of the galaxy, but also for the evolution of the Universe. This is why there are various theoretical models trying to describe the SFH of galaxies. One of those models is the Delayed-Tau model, which approximates the star formation of galaxies as a function with a rising Star Formation Rate (SFR) at the beginning, until it reaches a peak at a time $\tau$, unique for each galaxy, and then it drops at an exponential rate.

@@

**Keywords:** Galaxies, Galaxy Evolution, Star Formation History (SFH), Star Formation Rate (SFR), Delayed-$\tau$, Local Cosmological Volume, Lilly-Madau Plot, Redshift, Newton-Ramphson, Markov Chain Monte Carlo (MCMC).

## Greek

Οι γαλαξίες είναι τεράστια συστήματα από αστέρια, πλανήτες, αέρια και σκόνη που συγκρατούνται μαζί λόγω της βαρύτητας. Κάθε γαλαξίας αποτελεί ένα «κοσμικό νησί» με δισεκατομμύρια ή ακόμα και τρισεκατομμύρια αστέρια, πλανήτες, μαύρες τρύπες και άλλα αστρονομικά αντικείμενα. Κάθε γαλαξίας διαφέρει τόσο στο πόσο μαζικός είναι όσο και στην μορφολογία του. Συνεχώς, εντώς των γαλαξιών, νεφελώματα τα οποία αποτελούνται από αέρια και σκόνη, καταρρέουν λόγω της βαρύτητας τους και παράγουν αστέρια. Η παραγωγή των άστρων επιρρεάζεται από τα χαρακτηριστικά του εκάστοτε γαλαξία και μέσω της μελέτης της ιστορίας της αστρογέννεσης των γαλαξιών, μπορούμε να καταλάβουμε την προέλευση των αστρικών και πλανητικών συστημάτων αλλά και την εξέλιξη τόσο των γαλαξιών όσο και του σύμπαντος.

Μέσω φασματοσκοπικών και φωτομετρικών μεθόδων μπορούμε να υπολογίσουμε των ρυθμό αστρογέννεσης (Star Formation Rate, SFR) των γαλαξίων σε διάφορες ηλικίες του γαλαξία (διαφορετικά φάσματα που χρησιμοποιούνται για τον υπολογισμό διαφορετικών αστρικών πληθυσμών), και με αυτών των τρόπο είναι δυνατόν να μάθουμε την ιστορία αστρογένεσης του κάθε γαλαξία (Star Formation History, SFH). Κάποια μοντέλα προσπαθούν να παραμετροποιήσουν την SFH, χρησιμοποιούντας συναρτήσεις της μορφής $\text{SFR} = f(t)$, προσπαθώντας να προσεγγίσουν την πραγματική SFH με μία ομαλή και εξιδινακευμένη εξίσωση που αγνοεί την ποικιλομορφία των γαλαξιών.

@@

# Star Formation
ing professor
The SFH of a galaxy describesution of its star formation rate over time. By selecting an appropriate model for SFH, we can analyze stellar production, predict periods of active or quiescent star formation, and determine when SFR stabilizes.

@@
